---
title: "XGBoost - complete cytogenetic case analysis"
output: html_notebook
---

```{r}
library(tidymodels)
library(doParallel)

set.seed(122)
mcl_split <- initial_split(MCL, strata=status)
mcl_train <- training(mcl_split)
mcl_test <- testing(mcl_split)

mcl_folds<-vfold_cv(mcl_train, strata=status)

```



```{r}
mcl_recipe<-recipe(status ~.,data=mcl_train) %>% 
  update_role(pt_id, new_role = "id variable") %>% 
  step_other(all_nominal_predictors(),threshold = 0.01) %>% 
  step_dummy(all_nominal_predictors()) %>% 
  step_nzv(all_predictors()) 

   



mcl_prep <- prep(mcl_recipe)
mcl_data_preprocessed <- juice(mcl_prep)
juice(mcl_prep)
```

#Tune hyperparameters

```{r}
xgb_spec <- boost_tree(
  trees= 2000,
  tree_depth = tune(),
  min_n = tune(),
  loss_reduction = tune(),
  sample_size = tune(),
  mtry= tune(),
  learn_rate = tune()) %>% 
  set_engine("xgboost", ) %>% 
  set_mode("classification")

xgb_spec
```

#Use grid based / space fitting to cover hyperparameter space

```{r}
xgb_grid <- grid_latin_hypercube(
  min_n(),
  tree_depth(),
  loss_reduction (),
  learn_rate(),
  finalize(mtry(), mcl_train),
  sample_size = sample_prop(),
  size = 50
)



xgb_grid 
```

#Finalize workflow with recipe and hyperparameter tuning specifications

```{r}
xgb_wf <- workflow() %>%
  add_recipe(mcl_recipe) %>% 
  add_model(xgb_spec)

xgb_wf
```

#Make function for getting variable importance from folds

```{r}
get_xgb_imp <- function(x) {
  x %>% 
    extract_fit_parsnip() %>% 
    vip::vi()
}
```

#Set up parallel processing #Tune using the grid of hyperparameters and
the resampled folds #You may get warnings for new levels created for
sparse features - the hyperparameter fit will still work.

```{r}
doParallel::registerDoParallel()

set.seed(122)
xgb_res <- tune_grid(
  xgb_wf,
  resamples = mcl_folds,
  grid = xgb_grid,
  control = control_grid(extract=get_xgb_imp,save_pred = TRUE)
)

xgb_res
```

#Collect metrics

```{r}
collect_metrics(xgb_res)
```

#Visualize hyperparameter metrics for the possible models

```{r}
xgb_res %>%
  collect_metrics() %>%
  filter(.metric == "roc_auc") %>%
  select(mean, mtry:sample_size) %>%
  pivot_longer(mtry:sample_size,
               values_to = "value",
               names_to = "parameter"
  ) %>%
  ggplot(aes(value, mean, color = parameter)) +
  geom_point(alpha = 0.8, show.legend = FALSE) +
  facet_wrap(~parameter, scales = "free_x") +
  labs(x = NULL, y = "AUC")
```

#Show the best performing set of parameters

```{r}
best_models <- show_best(xgb_res, "roc_auc")
best_models

```

#Select best hyperparameters based on AUC

```{r}
best_auc <- select_best(xgb_res, "roc_auc")
best_auc
```

#Look at feature importance in the trained stratified folds.

```{r}
library(vip)
library(jcolors)

xgb_res %>% 
  select(id, .extracts) %>% 
  unnest(.extracts) %>% 
  unnest(.extracts) %>% 
  group_by(Variable) %>% 
  summarise(Mean=mean(Importance),
            Variance=sd(Importance)) %>% 
  slice_max(Mean, n=30) %>% 
  ggplot(aes(Mean, reorder(Variable, Mean))) +
  geom_crossbar(aes(xmin=Mean-Variance, xmax=Mean+Variance)) +
    labs(x="Variable Importance", y=NULL)


```

#Finalize workflow with the hyperparameter values

```{r}
final_xgb <- 
  finalize_workflow(xgb_wf, best_auc)

final_xgb
```

##Incorporate the hyperparameter values into the last fit and fit the
model on the test set.

```{r}
xgb_last_all <- final_xgb %>% 
  last_fit(mcl_split) %>% mutate(model="All Feature Types")
```

#Evaluate model performance.

```{r}
final_res <- last_fit(final_xgb, mcl_split)
final_metrics_all <- collect_metrics(final_res)
final_metrics_all
```

#Make a confusion matrix showing predictions.

```{r}
collect_predictions(xgb_last_all) %>% 
  conf_mat(status, .pred_class) 
```

#Which variables were most important to the prediction? Make Variable
Importance Plot (VIP).

```{r}
library(vip)
library(jcolors)
xgb_fit <-extract_fit_parsnip(final_res)
vip(xgb_fit, geom="col", num_features=20, aesthetics=list(color="grey50", fill="#1b98e0")) + scale_x_discrete(labels=c("ldh"="LDH", "platelets" ="Platelets", "ki_67"="Ki67%", "hgb"="Hgb","bm"= "BM", "b2m"="B2M", "wbc"="WBC","age_dx_years"="Age at dx", "tp53_yes"= "TP53 mutation", "ecog_X1"= "ECOG 1", "total_genes_mutated" = "Number of Somatic Mutations", "morph_classic" ="Classic Morphology", "bmi"= "BMI", "b_symptoms_yes"="B symptoms", "smoke_no"="Never Smoker", "alcohol" = "Alcohol consumption", "card11_yes" = "CARD11 mutation", "gi_involve_yes" = "GI Involvement", "complex_kar_yes" = "Complex Karyotype", "site_ln_multiple" = "Site: Multiple Lymph Nodes"))
                                                                            
                                                                          
```

#Get model agnostic Shapley additive explanations (SHAP) and plot.

```{r}
library(SHAPforxgboost)
mcl_shap <-
  shap.prep(
    xgb_model = extract_fit_engine(xgb_fit),
    X_train = bake(mcl_prep,
    has_role("predictor"),
    new_data=NULL,
    composition="matrix"
    ), top_n=20
  )


```

```{r}
shap.plot.summary(mcl_shap) + scale_x_discrete(labels=c("ldh"="LDH", "platelets" ="Platelets", "ki_67"="Ki67%", "hgb"="Hgb","bm"= "BM", "b2m"="B2M", "wbc"="WBC","age_dx_years"="Age at dx", "tp53_yes"= "TP53 mutation", "ecog_X1"= "ECOG 1", "total_genes_mutated" = "Number of Somatic mutations", "morph_classic" ="Classic Morphology", "bmi"= "BMI", "b_symptoms_yes"="B symptoms", "smoke_no"="Never Smoker", "alcohol" = "Alcohol consumption", "card11_yes" = "CARD11 mutation", "gi_involve_yes" = "GI Involvement", "complex_kar_yes" = "Complex Karyotype", "site_ln_multiple" = "Site: Multiple Lymph Nodes", "total_germ_variants" = "Number of Germline Mutations"), limits=rev) 
```

#Plot ROC

```{r}
final_res %>%
  collect_predictions() %>%
  group_by(id) %>% 
  roc_curve(status, .pred_aggressive) %>%
  ggplot(aes(x = 1 - specificity, y = sensitivity, color=id)) +
  geom_line(size = 1.5, color = "#1b98e0") +
  geom_abline(
    lty = 2, alpha = 0.5,
    color = "gray50",
    size = 1.2
  )
```

#Make a dataframe of final predictions for visualization

```{r}
all_feature_model_predict <- final_res %>%
  collect_predictions() %>%
  group_by(id)
```
